---
title: "Midterm Report Part 3"
output: html_document
---

## **Contributors:**
> William Loving (wfl9zy)
> James Sweat (jes9hd)

## **Goals:**
> - Our goal here in part 3 is to explore jobs and job satisfaction on a more granular level.
> - Find out whether or not older individuals favor non-travel working styles or overtime positions.
> - Find out if job role and distance from home have an impact on general job satisfaction.
> - Find out if job satisfaction declines as the number of years since promotion rises.
> - Find out if the marital status of individuals when working has a general trend, are more people married? Divorced? Single?

# **Part 3: Job Satisfaction Outside of CS/DS:**
> - This data more-so concerns people and how they felt about their jobs rather than data about the job itself.


```{r, message=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(plotly)
```

# **Load Data:**

```{r, message=FALSE}
data <- read_csv("../data/job-satisfaction-data/hr_data.csv.csv")
```

```{r}
head(data)
```


# **Make Some Transformations for Ease of Understanding:**

```{r}

data$JobSatisfaction <- as.factor(data$JobSatisfaction)

```



# **Create Visuals:**

# Plot 1: 

```{r, echo=FALSE}

plot <- ggplot(data, aes(x=BusinessTravel, y=JobRole, fill=MonthlyRate)) + 
    geom_tile() +
    labs(
    x="Travel Type",
    y="Job Role",
    title = "Travel type as it relates to Job Role and Monthly Salary",
    fill="Monthly Rate"
 ) + 
  theme_minimal() + 
  scale_fill_gradient(low="white", high="green")



plot
```

> - This visual allows us to discern some interesting things. For example, Managers that travel frequently tend to be paid less, the reverse is true for Manufacturing Directors that don't travel. 


# Plot 2:


```{r, echo=FALSE}
plot <- ggplot(data, aes(x=JobSatisfaction, y=JobRole, fill=Age)) + 
  geom_tile() + 
  labs(
    x="Job Satisfaction",
    y="Job Role",
    fill="Age",
    title="Job Satisfaction as it relates to Job Role and Age"
  ) +
  scale_fill_gradient(low="white", high="red") +
  theme_minimal()

plot
```

> - Here we can gleam some small correlations, perhaps young people simply aren't made for management, manufacturing, or researching positions.

# Plot 3: 


```{r, echo=FALSE}

plot <- ggplot(data, aes(x=YearsSinceLastPromotion, y=MonthlyRate, fill=JobSatisfaction)) + 
  geom_bar(stat='identity', position="dodge") + 
  #geom_point() +
  labs(
    x="Years Since Promotion",
    y="Monthly Rate",
    fill = "Job Satisfaction",
    title="Years Since Promotion as it Relates to Monthly Rate and Job Satisfaction"
  ) + 
  theme_minimal()


ggplotly(plot)
```

> - Here we can discern that as the years since promotion go on, the job satisfaction definitely wavers, especially if the monthly rate is low and the years since promotion is high.



# Plot 4:


```{r, echo=FALSE}

plot <- ggplot(data, aes(x=MaritalStatus, y=MonthlyIncome, fill=Gender)) + 
  geom_bar(stat='identity') + 
  labs(
    x="Marital Status",
    y="Monthly Income",
    title = "Do People Tend to be Married when Working?",
    fill="Gender"
 ) + 
  theme_minimal()


ggplotly(plot)


```


> - Here we are simply showing that most working class people tend to be married regardless of their gender, we have significantly more monthly rate data-points for married individuals than Divorced or Single.



# Plot 5:


```{r, echo=FALSE}

plot <- ggplot(data, aes(x=JobSatisfaction, y=JobRole, fill=MonthlyIncome)) + 
    geom_tile() +
    labs(
    x="Job Satisfaction",
    y="Job Role",
    title = "What Jobs Pay the Most?",
    fill="Monthly Income"
 ) + 
  theme_minimal() + 
  scale_fill_gradient(low="white", high="blue")


plot
```


