---
title: "Midterm Report Part 2"
author: "wlf9zy, jes9hd"
date: "2023-10-23"
output: html_document
---

## **Contributors:**
> William Loving (wfl9zy)
> James Sweat (jes9hd)

## **Goals:**
> 1. Explore and visualize the broader Software Engineering field
> 2. Create story-telling visualizations that provide sneak-peaks into the psyche of many comapnies as well as working individuals.
> 3. Learn how to develop meaningful visualizations to communicate the data we have to an uninformed audience.
  
# **Part 2: Software Engineering Positions:**
> - Here our goal will be to create as many interesting plots as possible, this dataset is more feature rich so we should be able to find some interesting things.
> - Note: This dataset was taken from job data in India, we hope the trends seen there will to some degree translate.

```{r, message=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(plotly)
```

# Load Data: 

```{r}
data <- read_csv("../data/india-data/Salary_Dataset_with_Extra_Features.csv")

head(data)

```

# Make Some Transformations for Ease of Understanding:

```{r}

#Remove the one outlier
threshold <- 9000000

filtered_data <- data%>%
  filter(Salary < threshold)

filtered_data$Location <- as.factor(filtered_data$Location)

sampled_data  <- filtered_data[sample(nrow(data), 1000), ]


```

# Create Visualizations: 

```{r}

plot <- ggplot(filtered_data, aes(x=Rating, y=Salary)) +
          geom_point() + 
          labs(
            x="Rating",
            y="Salary",
            title="Correlation of Rating and Job Salary"
          ) + 
          theme_minimal()

plot
```





```{r}

plot <- ggplot(sampled_data, aes(x=`Job Roles`
                         , y=`Location`
                         , fill=Rating
                         , text = paste("Location: ", Location, "<br>Job Role: ", `Job Roles`, "<br>Rating: ", Rating))) +
                          geom_tile() + 
                          labs(
                            x = "Job Roles",
                            y = "Location",
                            fill = "Rating",
                            title = "Heatmap of Jobs By Location With Ratings"
                          ) + 
                          scale_fill_gradient(low="white", high="blue") + 
                          theme_minimal()


plot
```



```{r}

plot <- plot_ly(sampled_data, x = ~Rating, y = ~Salary, z = ~`Salaries Reported`, color=~Location, type = "scatter3d", mode = "markers")
plot

```