---
title: "Midterm Report Part 3"
author: "wlf9zy, jes9hd"
date: "2023-10-23"
output: html_document
---
---
title: "Midterm Report Part 2"
author: "wlf9zy, jes9hd"
date: "2023-10-23"
output: html_document
---

## **Contributors:**
> William Loving (wfl9zy)
> James Sweat (jes9hd)

## **Goals:**
> 1. Explore and visualize the broader Software Engineering field
> 2. Create story-telling visualizations that provide sneak-peaks into the psyche of many comapnies as well as working individuals.
> 3. Learn how to develop meaningful visualizations to communicate the data we have to an uninformed audience.
  
# **Part 2: Software Engineering Positions:**
> - Here our goal will be to create as many interesting plots as possible, this dataset is more feature rich so we should be able to find some interesting things.


```{r, message=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(plotly)
```

# Load Data: 

```{r}
data <- read_csv("../data/data-analyst-data/DataAnalyst.csv")


head(data)
```


# Make Small Transformation Using Function
```{r}

# extracts the higher salary value from string like "$47K-$76K (Glassdoor est.)"
extract_higher_salary <- function(salary_string) {
  matches <- regmatches(salary_string, regexpr("\\$(\\d+)K-\\$(\\d+)K", salary_string))
  if (length(matches) > 0) {
    salary_values <- as.integer(unlist(matches))
    if (length(salary_values) == 2){
      return(salary_values)
    }
  } 
    return(NA)
  }


data <- data %>%
  mutate(Lower_Salary_K = sapply(`Salary Estimate`, function(x) extract_higher_salary(x)[1]),
         Higher_Salary_K = sapply(`Salary Estimate`, function(x) extract_higher_salary(x)[2]))


View(data)

```
