---
title: "Midterm Report Part 2"
author: "wlf9zy, jes9hd"
date: "2023-10-23"
output: html_document
---

## **Contributors:**
> William Loving (wfl9zy)
> James Sweat (jes9hd)

## **Goals:**
> - Find out if higher salary jobs tend to be rated more highly than lower salary. 
> - Find out where jobs are offered and see if some jobs are location specific.
> - 
  
# **Part 2: Software Engineering Positions in India:**
> - The data is vast so visualizations must be poignant and intentional with the correlations we hope to explore. 
> - We'll begin by loading the data and making small transformations for ease of use.

```{r, message=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(plotly)
```

# **Load Data:** 

```{r, message=FALSE}
data <- read_csv("../data/india-data/Salary_Dataset_with_Extra_Features.csv")
```

```{r}
head(data)

```

# **Make Some Transformations for Ease of Understanding:**

```{r, message=FALSE}

#Remove the one outlier
threshold <- 9000000

data_filtered <- data%>%
  filter(Salary < threshold)

data_filtered$Location <- as.factor(data_filtered$Location)
data_filtered$`Employment Status` <- as.factor(data_filtered$`Employment Status`)

```


# **Create Visuals:** 

# Plot 1:

```{r, echo=FALSE}

plot <- ggplot(data_filtered, aes(x=Rating, y=Salary)) +
          geom_point() + 
          labs(
            x="Rating",
            y="Salary",
            title="Correlation of Rating and Job Salary"
          ) + 
          theme_minimal()
plot
```

> - Here we can clearly see a positive correlation between rating and salary. A rise in salary tends to correspond with a rise in rating from those taking the jobs.
> - It is also worth noting that there are still jobs receiving 5-star ratings that have lower salaries, the trend is clearly there, but the correlation is not absolute.




# Plot 2: 

```{r,echo=FALSE}
plot <- ggplot(data_filtered
          , aes(x=`Job Roles`
             , y=`Location`
             , fill=Rating)) + 
            geom_tile() + 
            labs(
              x = "Job Roles",
              y = "Location",
              fill = "Rating",
              title = "Heatmap of Jobs By Location With Ratings"
            ) + 
            scale_fill_gradient(low="white", high="blue") + 
            theme_minimal() + 
            theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

plot
```

> - This plot allows us to find out where certain jobs are offered and how those jobs are rated by the employees. You can see where certain jobs are not offered as well as where the highest rated jobs can be found. If you want a great Mobile development experience, go to Bangalore, for example.


# Plot 3:

```{r,echo=FALSE}

plot <- ggplot(data_filtered, aes(x=`Job Roles`, y=`Employment Status`, fill=Salary)) + 
  geom_tile() + 
  theme_minimal() + 
  scale_fill_gradient(low="white", high="red") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

plot
```

> - Here we have a similar heat-map to the above plot, but we have Employment Status and Salary as our fill variable. This tells us where the highest paying jobs are by their Employment Status. 
> - An experienced Java developer would likely look to full-time java positions to maximize his/her salary expectation.