---
title: "Midterm Report Part-1"
author: "wlf9zy, jes9hd"
date: "2023-10-22"
output: html_document
---
## **Contributors:**
> William Loving (wfl9zy)
> James Sweat (jes9hd)

## **Goals:**
> 1. Explore and visualize the broader Computer Science / Data Analysis industry fields. 
> 2. Discover interesting correlations between attributes of available jobs using multiple different Datasets.
> 3. Learn how to develop meaningful visualizations to communicate the data we have to an uninformed audience.
  
# **Part 1: Data Science Positions:**
> Here we will explore Data Scientist Jobs in an around the United States
  
# Load Data:


```{r, echo=FALSE, message=FALSE}

library(readr)
library(dplyr)
library(ggplot2)
library(plotly)

```

```{r}
data <- read_csv("../data/data-science-jobs/ds_salaries.csv")

head(data)
```

# Explore Data and Make Necessary Transformations:

```{r, echo=FALSE}
str(data)
```

```{r, echo=FALSE}
summary(data)
```


## Mutate Categorical Variables to be More Descriptive: 

```{r, echo=TRUE}
data_transformed <- data%>%
  mutate(experience_level = ifelse(experience_level=="EN", "Entry-Level", 
                                   ifelse(experience_level=="MI", "Manager-Level", 
                                          ifelse(experience_level=="SE", "Senior-Level",
                                                 ifelse(experience_level=="EX", "Executive-Level", experience_level)))))

data_transformed <- data_transformed%>%
  mutate(employment_type = ifelse(employment_type=="CT", "Contract-Work", 
                                   ifelse(employment_type=="FT", "Full-Time", 
                                          ifelse(employment_type=="PT", "Part-Time",
                                                 ifelse(employment_type=="FL", "FreeLance", employment_type)))))

data_transformed <- data_transformed%>%
  mutate(company_size = ifelse(company_size=="L", "Large", 
                                   ifelse(company_size=="M", "Medium", 
                                          ifelse(company_size=="S", "Small", company_size))))

data_transformed <- data_transformed%>%
  mutate(remote_ratio = ifelse(remote_ratio==0, "In-Person", 
                                   ifelse(remote_ratio==50, "Hybrid", 
                                          ifelse(remote_ratio==100, "Remote", remote_ratio))))


head(data_transformed)

```

# **Create Plots to Tell Stories:**

## Plot 1:
> - With this plot we can clearly see that as your experience level rises, you can expect to see a corresponding increase in salary.
> - It is also worth noting that different types of work see different effects, for example, contract work is much more volatile than Full Time salaries.

```{r, echo=TRUE}
plot <- ggplot(data_transformed, aes(x=experience_level, y=salary_in_usd, fill=employment_type)) +
  geom_bar(stat='identity', position='dodge') + 
  labs(
    x="Experience Level",
    y="Salary in $USD",
    fill="Employment Type",
    title="The effects of Experience Level on Salary"
  ) + 
  scale_x_discrete(limits = c("Entry-Level", "Senior-Level", "Manager-Level","Executive-Level")) +
  theme_minimal()

ggplotly(plot)
```

## Plot 2:
> - Note that In-Person only paid the highest for Medium Sized Companies, Remote actually had the highest payout for Large 
> - Small companies pay grows step-wise with respect to the remote ratio (Hybrid->In-Person->Remote)

```{r}
plot <- ggplot(data_transformed, aes(x=company_size, y=salary_in_usd, fill=remote_ratio)) +
          geom_bar(stat='identity', position='dodge') +
          labs(
            x="Company size",
            y="Salary in $USD",
            fill="Remote Ratio",
            title="The effects of Company Size and Remote Ratio on Salary"
          ) + 
          scale_x_discrete(limits = c("Small", "Medium", "Large")) + 
          theme_minimal()

ggplotly(plot)


```

## Plot 3:
> - A lot of information, but the most interesting is that the US has the highest paying jobs by far with Small companies in Japan as a close second.

```{r}

plot <- ggplot(data_transformed, aes(x=company_location, y=salary_in_usd, fill=company_size)) +
          geom_bar(stat='identity', position='dodge') +
          labs(
            x="Company Location",
            y="Salary in $USD",
            fill="Company Size",
            title="The effects of Company Location and Size on Salary"
          ) + 
          theme_minimal() +
          theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))



  
  
ggplotly(plot)
```


